Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-08-16 prop_letter.tex
some work on KS

Our first result is that a QIVPM is consistent with the inability to
avoid uncertainty found in conventional infinite precise quantum
mechanics~\cite{PeresRon1988,peres1995quantum}.

The technical question is what happens if we relax the perfect
measurement assumption that the intervals are precisely $[0,0]$ and
$[1,1]$. We need to examine what happens if measurement imperfections
occasionally cause an ideally impossible event to be detected or an
ideally certain event not to be detected. Mermin argues in this case
that:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case~\citep{Mermin1999}.
\end{quote}

We can formally test this hypothesis by considering the notion of
a $\delta$-deterministic QIVPM below.

\begin{definition}\label{def:delta-deterministic}We define
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
  as $\delta$-deterministic if either
  $\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]$ or
  $\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]$ for any
  $P\in\events$.\end{definition}

As $\delta$ gets bigger, the QIVPM allows more and more indeterminate
behavior. A $\frac{1}{3}$-deterministic QIVPM can be given by
$\bar{\mu}\left(P\right)=\left[x/3,x/3\right]$, where $P$ projects
onto an $x$-dimensional subspace of a three-dimensional Hilbert space.

We now prove that for small enough $\delta$, the Kochen-Specker result is
still valid, i.e., if the measurement imperfections are small enough,
we cannot nullify the Kochen-Specker theorem as
Meyer~\cite{PhysRevLett.83.3751} argued.

\begin{thm}[Kochen-Specker~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}]\label{thm:Kochen-Specker}
  Given a Hilbert space $\Hilb$ of dimension~$d\ge3$, there is no
  0-deterministic
  measure~$\mu^{\mathrm{D}}:\events\rightarrow\left\{ 0,1\right\}
  $.\end{thm}

We will adapt the proof of the Kochen-Specker
theorem~\cite{kochenspecker1967,
  peres1995quantum,Redhead1987-REDINA,Griffiths2003} to show that it
is, in general, impossible to have just the two intervals \imposs\ and
\necess: every QIVPM must be parameterized by a collection of
intervals that includes at least one additional ``fuzzy'' interval and
some events must be associated with such uncertain intervals. The
significance of this observation is that the uncertainty in the
intervals is, in contrast to the classical case, \emph{a fundamental
  part of quantum theory}. There are in fact two sources of
uncertainty: one due to the finite measurement resources and one due
to complementarity, which requires every quantum framework to be truly
probabilistic.

We will proceed by assuming a 0-deterministic QIVPM $\bar{\mu}$ and
derive a contradiction using the following nine ``Mermin-Peres
square''
observables~$\mathbf{O}_{ij}$~\cite{Mermin1990Simple,peres1995quantum}:
\begin{center}
\begin{tabular}{|c|c|c|}
\hline 
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\mathbb{1}$  & $\sigma_{z}\otimes\sigma_{z}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{x}\otimes\sigma_{x}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$ \tabularnewline
\hline 
\end{tabular}\,,
\par\end{center}

\noindent where $\left\{\mathbb{1},\sigma_{x},\sigma_{y},\sigma_{z}\right\}$ are
Pauli matrices acting on a pair of spin-$\frac{1}{2}$ states. We now define the
expectation value of the QIVPM for our Mermin-Peres square.

\begin{definition}Given a Hilbert space $\Hilb$ of dimension~$d\ge3$,
  for every QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, any
  observable~$\mathbf{O}$ with spectral decomposition
  $\mathbf{O}=\sum_{i=1}^{N}\lambda_{i}P_{i}$, and the minimal
  subspace of events $\events'$ containing
  $\left\{ P_{i}\right\} _{i=1}^{N}$, we have
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\,.\label{eq:quantum-interval-expectation-core}
\end{equation}
\end{definition}

@@

@@

@@

Foundational concepts in quantum mechanics, such as the Kochen-Specker
and Gleason theorems, rely in a very subtle way on the use of
unbounded resources. By assuming infinitely precise measurements,
these two insightful theorems underlie two fundamental aspects of
quantum mechanics. On the one hand, Kochen-Specker's result reveals a
distinctive aspect of physical reality, the fact that it is
contextual, and, on the other hand, Gleason's theorem establishes a
relationship between quantum states and probabilities uniquely defined
by Born's rule.  Our goal in this paper has been to analyze the
physical consequences of a mathematical framework that allows for
finite precision measurements by introducing the concept of quantum
interval-valued probability. This framework incorporates uncertainty
in the measurement results by defining fuzzy probability measures, and
includes standard quantum measurement as the particular instance of
sharp, infinitely precise, intervals.  In addition, we showed how
these two theorems emerge as limiting cases of this same framework,
thus connecting two seemingly unrelated aspects of quantum physics. We
noted that {\it arbitrary\/} finite precision measurement conditions
can nullify the main tenets of both theorems. However, by carefully
specifying experimental uncertainties, we were able to establish
rigorous bounds on the validity of these two theorems. Therefore, we
have established a context in which infinite precision quantum
mechanical theories can be reconciled with finite precision quantum
mechanical measurements, and have provided a possible resolution of
the Meyer-Mermin debate on the impact of finite precision on the
Kochen-Specker theorem \cite{PhysRevLett.83.3751,Mermin1999}.

>>>>

Our generalization of quantum probability measures to QIVPMs will
allow us to strengthen one of the fundamental theorems in of quantum
information: the Kochen-Specker
theorem~\cite{PeresRon1988,peres1995quantum}. Our stronger version of
the theorem will suggest a resolution to the debate initiated by Meyer
and Mermin on the relevance of the Kochen-Specker to experimental, and
hence finite-precision, quantum
measurements~\citep{PhysRevLett.83.3751,Mermin1999,Kent1999,HavlicekKrennSummhammerSvozil2001,SimonBruknerZeilinger2001,Cabello2002,Larsson2002,Appleby2002,Peres2003,BarrettKent2004,Appleby_2005,Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.
. Specifically, the original Kochen-Specker theorem is formulated
under the assumption that all observables of a quantum mechanical
system have \emph{definite values at all
  times}~\cite{StanfordXXX}. Our interval-valued framework will allow
us to state, and prove, a stronger version of the theorem that holds
even if the observables have values that are only definite up to some
specific parameter that quantifies experimental uncertainty. Put
differently, our stronger statement and theorem quantifies Mermin's
qualitative argument:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case~\citep{Mermin1999}.
\end{quote}

\paragraph*{$\delta$-Determinism.} The first step in our formalization
is to introduce a family of QIVPMs parameterized by an uncertainty
$\delta$ which we call $\delta$-deterministic QIVPMs.

\begin{definition}\label{def:delta-deterministic} A
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is
  $\delta$-deterministic if, for every event $P\in\events$, we have
  that either 
  $\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]$ or
  $\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]$. 
\end{definition}

When $\delta=0$ , every event must be assigned a probability in either
the interval $[0,0]$ or $[1,1]$, i.e., every event is completely
determined with certainty. As $\delta$ gets bigger, the QIVPM allows
more and more indeterminate behavior. Our technical result will be
that for ``small enough'' $\delta$, a variant of the Kochen-Specker
theorem is still valid. To state and prove this result we will first
reformulate the original Kochen-Specker as a statement regarding the
non-existence of $0$-deterministic QIVPM. We will then strengthen this
result to show the non-existence of any $\delta$-deterministic QIVPM
when $\delta < 1/3$ and the Hilbert space dimension is greater than
2.

\paragraph*{Expectation Values of Fuzzy Observables.}  Our following
development will crucially rely on analyzing the expectation values of
families of observables. We introduce the necessary definitions before
stating and proving our variants of the Kochen-Specker theorem.

\begin{definition} Given a Hilbert space $\Hilb$ of dimension~$d\ge3$,
  for every QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, any
  observable~$\mathbf{O}$ with spectral decomposition
  $\mathbf{O}=\sum_{i=1}^{N}\lambda_{i}P_{i}$, and the minimal
  subspace of events $\events'$ containing
  $\left\{ P_{i}\right\} _{i=1}^{N}$, we have
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\,.\label{eq:quantum-interval-expectation-core}
\end{equation}
\end{definition}

@@

We now state and prove a theorem that is equivalent to the original
Kochen-Specker theorem but that is formulated in terms of
$0$-deterministic QIVPMs.

\begin{thm}[0-deterministic variant of Kochen-Specker] 
\label{thm:Kochen-Specker}
Given a Hilbert space $\Hilb$ of dimension~$d\ge3$, there is no
0-deterministic
measure~$\mu^{\mathrm{D}}:\events\rightarrow\left\{ 0,1\right\}$.
\end{thm}

To prove this result, we proceed by assuming a 0-deterministic QIVPM
$\bar{\mu}$ and derive a contradiction using the same nine observables
$\mathbf{O}_{ij}$ used in the ``Mermin-Peres'' proof of the
Kochen-Specker theorem~\cite{Mermin1990Simple,peres1995quantum}:
{\renewcommand{\arraystretch}{2}%
\begin{center}
\begin{tabular}{|@{\quad}c@{\quad}|@{\quad}c@{\quad}|@{\quad}c@{\quad}|}
\hline 
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\mathbb{1}$  & $\sigma_{z}\otimes\sigma_{z}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{x}\otimes\sigma_{x}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$ \tabularnewline
\hline 
\end{tabular}\,,
\par\end{center}
}
\noindent where $\left\{\mathbb{1},\sigma_{x},\sigma_{y},\sigma_{z}\right\}$ are
Pauli matrices acting on a pair of spin-$\frac{1}{2}$ states. We now define the
expectation value of the QIVPM for our Mermin-Peres square.

@@

We will now adapt the proof above to show that it is, in general,
impossible to have just the two intervals \imposs\ and \necess: every
QIVPM must be parameterized by a collection of intervals that includes
at least one additional ``fuzzy'' interval and some events must be
associated with such uncertain intervals. The significance of this
observation is that the uncertainty in the intervals is, in contrast
to the classical case, \emph{a fundamental part of quantum
  theory}. There are in fact two sources of uncertainty: one due to
the finite measurement resources and one due to complementarity, which
requires every quantum framework to be truly probabilistic.

???~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA,Griffiths2003}

@@

Foundational concepts in quantum mechanics, such as the Kochen-Specker
and Gleason theorems, rely in a very subtle way on the use of
unbounded resources. By assuming infinitely precise measurements,
these two insightful theorems underlie two fundamental aspects of
quantum mechanics. On the one hand, Kochen-Specker's result reveals a
distinctive aspect of physical reality, the fact that it is
contextual, and, on the other hand, Gleason's theorem establishes a
relationship between quantum states and probabilities uniquely defined
by Born's rule.  Our goal in this paper has been to analyze the
physical consequences of a mathematical framework that allows for
finite precision measurements by introducing the concept of quantum
interval-valued probability. This framework incorporates uncertainty
in the measurement results by defining fuzzy probability measures, and
includes standard quantum measurement as the particular instance of
sharp, infinitely precise, intervals.  In addition, we showed how
these two theorems emerge as limiting cases of this same framework,
thus connecting two seemingly unrelated aspects of quantum physics. We
noted that {\it arbitrary\/} finite precision measurement conditions
can nullify the main tenets of both theorems. However, by carefully
specifying experimental uncertainties, we were able to establish
rigorous bounds on the validity of these two theorems. Therefore, we
have established a context in which infinite precision quantum
mechanical theories can be reconciled with finite precision quantum
mechanical measurements, and have provided a possible resolution of
the Meyer-Mermin debate on the impact of finite precision on the
Kochen-Specker theorem~\cite{PhysRevLett.83.3751,Mermin1999}.
